\label{Cables}
\section*{Cables}
\index{Cables }
\addcontentsline{toc}{section}{Cables}

\label{index-Cables-191}
You can either use the cable that came with your UPS (the easiest if we
support it) or you can make your own cable. We recommend that you obtain a
supported cable directly from APC.  

If you already have an APC cable, you can determine what kind it is by
examining the flat sides of the two connectors where you will find the cable
number embossed into the plastic. It is generally on one side of the male
connector.  

To make your own cable you must first know whether you have a UPS that speaks
the apcsmart protocol or a ``dumb'' UPS that uses serial port line voltage
signalling.  

If you have an apcmart UPS, and you build your own cable, build a {\it
Smart-Custom} cable. If you have a voltage-Signalling or dumb UPS, build a
{\it Simple-Custom} cable. If you have a BackUPS CS with a RJ45 connector, you
can build your own {\it Custom-RJ45} cable. 

\label{Smart_002dCustom-Cable-for-SmartUPSes}

\subsection*{Smart-Custom Cable for SmartUPSes}
\index{SmartUPSes!Smart-Custom Cable for }
\index{Smart-Custom Cable for SmartUPSes }
\addcontentsline{toc}{subsection}{Smart-Custom Cable for SmartUPSes}

\label{index-Smart-Custom-Cable-192}
\label{index-Cables_002c-Smart-Custom-193}

\footnotesize
\begin{verbatim}
     
       SMART-CUSTOM CABLE
     
     Signal Computer                  UPS
            DB9F                     DB9M
      RxD    2   --------------------  2  TxD  Send
      TxD    3   --------------------  1  RxD  Receive
      GND    5   --------------------  9  Ground
\end{verbatim}
\normalsize

When using this cable with apcupsd specify the following in apcupsd.conf:  

If you have an OS that requires DCD or RTS to be set before you can receive
input, you might try building the standard APC Smart 940-0024C cable listed
below. 

\footnotesize
\begin{verbatim}
     
     UPSCABLE smart
     UPSTYPE apcsmart
     DEVICE /dev/ttyS0 (or whatever your serial port is)
\end{verbatim}
\normalsize

If you wish to build the standard cable furnished by APC (940-0024C), use the
following diagram. 

\footnotesize
\begin{verbatim}
     
       APC Smart Cable 940-0024C
     
     Signal Computer                  UPS
            DB9F                     DB9M
      RxD    2   --------------------  2  TxD  Send
      TxD    3   --------------------  1  RxD  Receive
      DCD    1   --*
                   |
      DTR    4   --*
      GND    5   --------------------  9  Ground
      RTS    7   --*
                   |
      CTS    8   --*
     
\end{verbatim}
\normalsize

\label{Smart-Signalling-Cable-for-BackUPS-CS-Models}

\subsection*{Smart Signalling Cable for BackUPS CS Models}
\index{Models!Smart Signalling Cable for BackUPS CS }
\index{Smart Signalling Cable for BackUPS CS Models }
\addcontentsline{toc}{subsection}{Smart Signalling Cable for BackUPS CS
Models}

\label{index-Cables_002c-Smart-Signalling-194}
\label{index-Cables_002c-BackUPS-CS-195}
If you have a BackUPS CS, you are probably either using it with the USB cable
that is supplied or with the 940-0128A supplied by APC, which permits running
the UPS in dumb mode. By building your own cable, you can now run the BackUPS
CS models (and perhaps also the ES models) using smart signalling and have all
the same information that is available as running it in USB mode.  

The jack in the UPS is actually a 10 pin RJ45. However, you can just as easily
use a 8 pin RJ45 connector, which is more standard (ethernet TX, and ISDN
connector). It is easy to construct the cable by cutting off one end of a
standard RJ45-8 ethernet cable and wiring the other end (three wires) into a
standard DB9F female serial port connector.  

Below, you will find a diagram for the CUSTOM-RJ45 cable: 

\footnotesize
\begin{verbatim}
     
       CUSTOM-RJ45 CABLE
     
     Signal Computer              UPS     UPS
            DB9F                 RJ45-8  RJ45-10
      RxD    2   ----------------  1      2     TxD  Send
      TxD    3   ----------------  7      8     RxD  Receive
      GND    5   ----------------  6      7     Ground
      FG  Shield ----------------  3      4     Frame Ground
     
     The RJ45-8 pins are: looking at the end of the connector:
     
      8 7 6 5 4 3 2 1
     ___________________
     | . . . . . . . . |
     |                 |
     -------------------
            |____|
     
     The RJ45-10  pins are: looking at the end of the connector:
     
     10 9 8 7 6 5 4 3 2 1
     _______________________
     | . . . . . . . . . . |
     |                     |
     -----------------------
            |____|
     
\end{verbatim}
\normalsize

For the serial port DB9F connector, the pin numbers are stamped in the plastic
near each pin. In addition, there is a diagram near the end of this chapter.  

Note, one user, Martin, has found that if the shield is not connected to the
Frame Ground in the above diagram (not in our original schematic), the UPS (a
BackUPS CS 500 EI) will be unstable and likely to rapidly switch from power to
batteries (i.e. chatter).  

When using this cable with apcupsd specify the following in apcupsd.conf: 

\footnotesize
\begin{verbatim}
     
     UPSCABLE smart
     UPSTYPE apcsmart
     DEVICE /dev/ttyS0 (or whatever your serial port is)
\end{verbatim}
\normalsize

The information for constructing this cable was discovered and transmitted to
us by slither\_man. Many thanks! 

\label{Voltage_002dSignalling-Cable-for-_0022dumb_0022-UPSes}

\subsection*{Voltage-Signalling Cable for ``dumb'' UPSes}
\index{UPSes!Voltage-Signalling Cable for "dumb" }
\index{Voltage-Signalling Cable for "dumb" UPSes }
\addcontentsline{toc}{subsection}{Voltage-Signalling Cable for "dumb"
UPSes}

\label{index-Cables_002c-dumb-196}
\label{index-dumb_002c-Cables-197}
{\bf NOTE. YOU DO NOT HAVE THIS CABLE UNLESS YOU BUILT IT YOURSELF.  THE
SIMPLE-CUSTOM CABLE IS NOT AN APC PRODUCT.}  

For ``dumb'' UPSes using voltage signalling, if you are going to build your
own cable, we recommend to make the cable designed by the apcupsd team as
follows: 

\footnotesize
\begin{verbatim}
     
            SIMPLE-CUSTOM CABLE
     
     Signal Computer                  UPS
            DB9F   4.7K ohm          DB9M
      DTR    4   --[####]--*              DTR set to +5V by Apcupsd
                           |
      CTS    8   ----------*---------  5  Low Battery
      GND    5   --------------------  4  Ground
      DCD    1   --------------------  2  On Battery
      RTS    7   --------------------  1  Kill UPS Power
     
\end{verbatim}
\normalsize

List of components one needs to make the Simple cable:  

\begin{enumerate}
\item One (1) male DB9 connector, use solder type connector only.  
\item One (1) female DB9/25F connector, use solder type connector only.  
\item One (1) 4.7K ohm 1/4 watt 5\% resistor.  
\item resin core solder.  
\item three (3) to five (5) feet of 22AWG multi-stranded four or more
   conductor cable.  
   \end{enumerate}

\begin{enumerate}
\item Solder the resistor into pin 4 of the female DB9 connector.  
\item Next bend the resistor so that it connects to pin 8 of the female DB9
   connector.  
\item Pin 8 on the female connector is also wired to pin 5 on the male DB9
   connector. Solder both ends.  
\item Solder the other pins, pin 5 on the female DB9 to pin 4 on the male
   connector; pin 1 on the female connector to pin 2 on the male connector; and
   pin 7 on the female connector to pin 1 on the male connector.  
\item Double check your work.  
   \end{enumerate}

We use the DTR (pin 4 on the female connector) as our +5 volts power for the
circuit. It is used as the Vcc pull-up voltage for testing the outputs on any
``UPS by APC'' in voltage-signalling mode.  This cable may not work on a
BackUPS Pro if the default communications are in apcsmart mode. This cable is
also valid for ``ShareUPS'' BASIC Port mode and is also reported to work on
SmartUPSes. However, the Smart Cable described above is much simpler. To have
a better idea of what is going on inside apcupsd, for the SIMPLE cable apcupsd
reads three signals and sets three: 

\footnotesize
\begin{verbatim}
     
     Reads:
     CD, which apcupsd uses for the On Battery signal when high.
     
     CTS, which apcupsd uses for the Battery Low signal when high.
     
     RxD (SR), which apcupsd uses for the Line Down
         signal when high. This signal isn't used for much.
     
     Sets:
     DTR, which apcupsd sets when it detects a power failure (generally
          5 to 10 seconds after the CD signal goes high). It
          clears this signal if the CD signal subsequently goes low
          -- i.e. power is restored.
     
     TxD (ST), which apcupsd clears when it detects that the CD signal
          has gone low after having gone high - i.e. power is restored.
     
     RTS, which apcupsd sets for the killpower signal -- to cause the UPS
          to shut off the power.
\end{verbatim}
\normalsize

Please note that these actions apply only to the SIMPLE cable, the signals
used on the other cables are different.  

Finally, here is another way of looking at the CUSTOM-SIMPLE cable: 

\footnotesize
\begin{verbatim}
     
     APCUPSD SIMPLE-CUSTOM CABLE
     
     Computer Side  |  Description of Cable           |     UPS Side
     DB9f  |  DB25f |                                 |   DB9m  | DB25m
     4     |   20   |  DTR (5vcc)             *below  |    n/c  |
     8     |    5   |  CTS (low battery)      *below  | <-  5   |   7
     2     |    3   |  RxD (no line voltage)  *below  | <-  3   |   2
     5     |    7   |  Ground (Signal)                |     4   |  20
     1     |    8   |  CD (on battery from UPS)       | <-  2   |   3
     7     |    4   |  RTS (kill UPS power)           | ->  1   |   8
     n/c   |    1   |  Frame/Case Gnd (optional)      |     9   |  22
     
     Note: the <- and -> indicate the signal direction.
     
     Optional connections of original SIMPLE-CUSTOM specification
     that are not used.
     
                   4.7K ohm
      DTR    4   --[####]--*              Note needed
                           |
      RxD    2   ----------*---------  3  Not used by Apcupsd
     
\end{verbatim}
\normalsize

When using this cable with apcupsd specify the following in apcupsd.conf: 

\footnotesize
\begin{verbatim}
     
     UPSCABLE simple
     UPSTYPE dumb
     DEVICE /dev/ttyS0 (or whatever your serial port is)
\end{verbatim}
\normalsize

\label{Other-APC-Cables-that-apcupsd-Supports}

\subsection*{Other APC Cables that apcupsd Supports}
\index{Other APC Cables that apcupsd Supports }
\index{Supports!Other APC Cables that apcupsd }
\addcontentsline{toc}{subsection}{Other APC Cables that apcupsd Supports}

apcupsd will also support the following off the shelf cables that are supplied
by APC  

\begin{itemize}
\item 940-0020B/C Simple Signal Only, all models.  
\item 940-0023A Simple Signal Only, all models.  
\item 940-0119A Simple Signal Only, Back-UPS Office, and BackUPS ES.  
\item 940-0024[B/C/G] SmartMode Only, SU and BKPro only.  
\item 940-0095[A/B/C] PnP (Plug and Play), all models.  
\item 940-1524C SmartMode Only  
\item 940-0127A/B USB Cables  
\item 940-0128A Simple Signal Only, Back-UPS CS in serial mode. 
   \end{itemize}

\label{Voltage-Signalling-Features-Supported-by-Apcupsd-for-Various-Cables}

\subsection*{Voltage Signalling Features Supported by Apcupsd for Various
Cables}
\index{Cables!Voltage Signalling Features Supported by Apcupsd for Various }
\index{Voltage Signalling Features Supported by Apcupsd for Various Cables }
\addcontentsline{toc}{subsection}{Voltage Signalling Features Supported by
Apcupsd for Various Cables}

The following table shows the features supported by the current version of
apcupsd (3.8.5 or later) for various cables running the UPS in
voltage-signalling mode.  

\addcontentsline{lot}{table}{Supported Features}
\begin{longtable}{p{1.2in}p{0.8in}p{1.2in}p{0.8in}p{1.0in}}
{Cable} & {Power Loss} & {Low Battery} & {Kill Power} & {Cable Disconnected 
 } \\
{940-0020B} & {Yes} & {No} & {Yes} & {No 
 } \\
{940-0020C} & {Yes} & {Yes} & {Yes} & {No 
 } \\
{940-0023A} & {Yes} & {No} & {No} & {No 
 } \\
{940-0119A} & {Yes} & {Yes} & {Yes} & {No 
 } \\
{940-0127A} & {Yes} & {Yes} & {Yes} & {No 
 } \\
{940-0128A} & {Yes} & {Yes} & {Yes} & {No 
 } \\
{940-0095A/B/C} & {Yes} & {Yes} & {Yes} & {No 
 } \\
{simple} & {Yes} & {Yes} & {Yes} & {No  
}

\end{longtable}

\label{Voltage-Signalling}

\subsection*{Voltage Signalling}
\index{Signalling!Voltage }
\index{Voltage Signalling }
\addcontentsline{toc}{subsection}{Voltage Signalling}

\label{index-Voltage-Signalling_002c-Cables-198}
\label{index-Cables_002c-Voltage-Signalling-199}
Apparently, all APC voltage-signalling UPSes have the same signals on the
output pins of the UPS. The difference at the computer end is due to different
cable configurations. Thus, by measuring the connectivity of a cable, one can
determine how to program the UPS. This is to be verified.  

The signals presented or accepted by the UPS on its DB9 connector using the
numbering scheme listed above is: 

\footnotesize
\begin{verbatim}
     
     UPS Pin         Signal meaning
      1     <-     Shutdown when set by computer for 1-5 seconds.
      2     ->     On battery power (this signal is normally low but
                         goes high when the UPS switches to batteries).
      3     ->     Mains down (line fail) See Note 1 below.
      5     ->     Low battery. See Note 1 below.
      6     ->     Inverse of mains down signal. See Note 2 below.
      7     <-     Turn on/off power (only on advanced UPSes only)
     
      Note 1: these two lines are normally open, but close when the
          appropriate signal is triggered. In fact, they are open collector
          outputs which are rated for a maximum of +40VDC and 25 mA. Thus
          the 4.7K ohm resistor used in the Custom Simple cable works
          quite well.
     
      Note 2: the same as note 1 except that the line is normally closed,
          and opens when the line voltage fails.
\end{verbatim}
\normalsize

\label{The-Back_002dUPS-Office-500-signals}

\subsection*{The Back-UPS Office 500 signals}
\index{Back-UPS Office 500 signals }
\index{Signals!Back-UPS Office 500 }
\addcontentsline{toc}{subsection}{Back-UPS Office 500 signals}

\label{index-Cables_002c-BackUPS-Office-200}
\label{index-BackUPS-Office_002c-cables-201}
The Back-UPS Office UPS has a telephone type jack as output, which looks like
the following: 

\footnotesize
\begin{verbatim}
     
     Looking at the end of the connector:
     
        6 5 4 3 2 1
       _____________
      | . . . . . . |
      |             |
      |  |----------|
      |__|
     
\end{verbatim}
\normalsize

It appears that the signals work as follows: 

\footnotesize
\begin{verbatim}
     
       UPS            Signal meaning
     1 (brown)    <-   Shutdown when set by computer for 1-5 seconds.
     2 (black)    ->   On battery power
     3 (blue)     ->   Low battery
     4 (red)           Signal ground
     5 (yellow)   <-   Begin signalling on other pins
     6 (none)          none
\end{verbatim}
\normalsize

\label{Analyses-of-APC-Cables}

\subsection*{Analyses of APC Cables}
\index{Cables!Analyses of APC }
\index{Analyses of APC Cables }
\addcontentsline{toc}{subsection}{Analyses of APC Cables}

\label{index-Cables-202}

\label{940_002d0020B-Cable-Wiring}

\subsubsection*{940-0020B Cable Wiring:}

\label{index-Cables-203}
This diagram is for informational purposes and is not complete. Although we do
not know what the black box semi-conductor contains, we believe that we
understand its operation (many thanks to Lazar M. Fleysher for working this
out).  

This cable can only be used on voltage-signalling UPSes, and provides the On
Battery signal as well as kill UPS power. Most recent evidence (Lazar's
analysis) indicates that this cable under the right conditions may provide the
Low Battery signal. This is to be confirmed. 

\footnotesize
\begin{verbatim}
     
     APC Part# - 940-0020B
     
     Signal Computer                  UPS
            DB9F                     DB9M
      CTS    8   --------------------  2  On Battery
      DTR    4   --------------------  1  Kill power
      GND    5   ---------------*----  4  Ground
                                |
                     ---        *----  9  Common
      DCD    1  ----|///|-----------   5  Low Battery
                    |\\\|
      RTS    7  ----|///| (probably a
                     ---   semi-conductor)
\end{verbatim}
\normalsize

Thanks to Lazar M. Fleysher. 

\label{940_002d0020C-Cable-Wiring}

\subsubsection*{940-0020C Cable Wiring:}

\label{index-Cables-204}
This diagram is for informational purposes and may not be complete, we don't
recommend that use it to build you build one yourself.  This cable can only be
used on voltage-signalling UPSes, and provides the On Battery signal, the Low
Battery signal as well as kill UPS power. In apcupsd versions 3.8.2 and prior,
please set your UPSCABLE to 940-0020B. In version 3.8.3 and later, you may
specify the cable as 940-0020C. Please note that this diagram may not be
accurate. 

\footnotesize
\begin{verbatim}
     
     APC Part# - 940-0020C
     
     Signal Computer                  UPS
            DB9F                     DB9M
      CTS    8   --------------------  2  On Battery
      DTR    4   --------------------  1  Kill power
      GND    5   ---------------*----  4  Ground
                                |
                                *----  9  Common
      RTS    7 -----[ 93.5K ohm ]----- 5  Low Battery
                    or semi-conductor
\end{verbatim}
\normalsize

\label{940_002d0023A-Cable-Wiring}

\subsubsection*{940-0023A Cable Wiring:}

\label{index-Cables-205}
This diagram is for informational purposes and may not be complete, we don't
recommend that use it to build you build one yourself.  This cable can only be
used on voltage-signalling UPSes, and apparently only provides the On Battery
signal. As a consequence, this cable is pretty much useless, and we recommend
that you find a better cable because all APC UPSes support more than just On
Battery. Please note that we are not sure the following diagram is correct. 

\footnotesize
\begin{verbatim}
     
     APC Part# - 940-0023A
     
     Signal Computer                  UPS
            DB9F                     DB9M
      DCD    1   --------------------  2  On Battery
     
                   3.3K ohm
      TxD    3   --[####]-*
                          |
      DTR    4   ---------*
      GND    5   ---------------*----  4  Ground
                                |
                                *----  9  Common
     
\end{verbatim}
\normalsize

\label{940_002d0095A-Cable-Wiring}

\subsubsection*{940-0095A Cable Wiring:}

\label{index-Cables-206}
This is the definitive wiring diagram for the 940-0095A cable submitted by
Chris Hanson \lt{}cph at zurich.ai.mit.edu\gt{}, who disassembled the original
cable, destroying it in the process. He then built one from his diagram and it
works perfectly. 

\footnotesize
\begin{verbatim}
     
     Construction and operation of the APC #940-0095A cable.
     This cable is included with the APC Back-UPS Pro PNP series.
     
     
     UPS end                                      Computer end
     -------                                      ------------
                       47k        47k
     BATTERY-LOW (5) >----R1----*----R2----*----< DTR,DSR,CTS (4,6,8)
                              |          |
                              |          |
                              |         /  E
                              |       |/
                              |    B  |
                              *-------|  2N3906 PNP
                                      |
                                      |\
                                        \  C
                                         |
                                         |
                                         *----< DCD (1)     Low Batt
                                         |
                                         |
                                         R 4.7k
                                         3
                                         |
                                  4.7k   |
     SHUTDOWN (1)    >----------*----R4----*----< TxD (3)
                              |
                              |  1N4148
                              *----K|---------< RTS (7)      Shutdown
     
     POWER-FAIL (2)  >--------------------------< RxD,RI (2,9) On Batt
     
     GROUND (4,9)    >--------------------------< GND (5)
     
     
     Operation:
     
     * DTR is "cable power" and must be held at SPACE.  DSR or CTS may be
     used as a loopback input to determine if the cable is plugged in.
     
     * DCD is the "battery low" signal to the computer.  A SPACE on this
     line means the battery is low.  This is signalled by BATTERY-LOW
     being pulled down (it is probably open circuit normally).
     
     Normally, the transistor is turned off, and DCD is held at the MARK
     voltage by TxD.  When BATTERY-LOW is pulled down, the voltage
     divider R2/R1 biases the transistor so that it is turned on, causing
     DCD to be pulled up to the SPACE voltage.
     
     * TxD must be held at MARK; this is the default state when no data is
     being transmitted.  This sets the default bias for both DCD and
     SHUTDOWN.  If this line is an open circuit, then when BATTERY-LOW is
     signalled, SHUTDOWN will be automatically signalled; this would be
     true if the cable were plugged in to the UPS and not the computer,
     or if the computer were turned off.
     
     * RTS is the "shutdown" signal from the computer.  A SPACE on this
     line tells the UPS to shut down.
     
     * RxD and RI are both the "power-fail" signals to the computer.  A
     MARK on this line means the power has failed.
     
     * SPACE is a positive voltage, typically +12V.  MARK is a negative
     voltage, typically -12V.  Linux appears to translate SPACE to a 1
     and MARK to a 0.
\end{verbatim}
\normalsize

\label{940_002d0095B-Cable-Wiring}

\subsubsection*{940-0095B Cable Wiring:}

\label{index-Cables-207}
This diagram is for informational purposes and may not be complete, we don't
recommend that use it to build one yourself. 

\footnotesize
\begin{verbatim}
     
     APC Part# - 940-0095B
     
     Signal Computer                  UPS
            DB9F                     DB9M
      DTR    4   ----*
      CTS    8   ----|
      DSR    6   ----|
      DCD    1   ----*
      GND    5   ---------------*----  4  Ground
                                |
                                *----  9  Common
      RI     9   ----*
                     |
      RxD    2   ----*---------------  2  On Battery
      TxD    3   ----------[####]----  1  Kill UPS Power
                           4.7K ohm
\end{verbatim}
\normalsize

\label{940_002d0119A-Cable-Wiring}

\subsubsection*{940-0119A Cable Wiring:}

\label{index-Cables-208}
This diagram is for informational purposes and may not be complete, we don't
recommend that use it to build you build one yourself. This cable is used with
the BackUPS Office UPSes. 

\footnotesize
\begin{verbatim}
     
     APC Part# - 940-0119A
     
       UPS      Computer
       pins     pins      Signal             Signal meaning
     1 (brown)    4,6      DSR DTR     <-   Shutdown when set by computer for 1-5 seconds.
     2 (black)    8,9      RI  CTS     ->   On battery power
     3 (blue)     1,2      CD  RxD     ->   Low battery
     4 (red)       5       Ground
     5 (yellow)    7       RTS         <-   Begin signalling on other pins
     6 (none)     none
     
\end{verbatim}
\normalsize

\label{BackOffice-ES}

\subsubsection*{BackOffice ES:}

\label{index-Cables-209}
\label{index-BackOffice-ES-210}
The BackUPS ES has a straight through serial cable with no identification on
the plugs. To make it work with apcupsd, specify the {\bf UPSCABLE 940-0119A}
and {\bf UPSTYPE backups}.  The equivalent of cable 940-0119A is done on a PCB
inside the unit.  Thanks to William Stock for supplying us with the
information about the straight through cable, the PCB, and the following
diagram: 

\footnotesize
\begin{verbatim}
     
     computer           ----------- BackUPS-ES -----------------
     DB9-M              DB-9F
     pin    signal      pin
     
      4      DSR   ->    4 --+
                             |  diode   resistor
      6      DTR   ->    6 --+---->|----/\/\/\---o kill power
     
      1      DCD   <-    1 --+
                             |
      2      RxD   <-    2 --+----------------+--o low battery
                                              |
      7      RTS   ->    7 --------+--/\/\/\--+
                                   |
                                   +--/\/\/\--+
                                              |
      8      RI    <-    8 --+----------------+--o on battery
                             |
      9      CTS   <-    9 --+
     
      5      GND   ---   5 ----------------------o ground
     
      3      TxD         3 nc
     
\end{verbatim}
\normalsize

\label{BackUPS-ES-and-CS-in-Serial-mode-with-Cable-940_002d0128A}

\subsubsection*{BackUPS ES and CS in Serial mode with Cable 940-0128A:}

\label{index-BackUPS-ES-211}
\label{index-Cables-212}
Though these UPSes are USB UPSes, APC supplies a serial cable (typically with
a green DB9 F connector) that has 940-0128A stamped into one side of the
plastic serial port connector. The other end of the cable is a 10 pin RJ45
connector that plugs into the UPS (thanks to Dean Waldow for sending me a
cable!). Apcupsd version 3.8.5 and later supports this cable when specified as
{\bf UPSCABLE 940-0128A} and {\bf UPSTYPE backups}. However, running in this
mode much of the information that would be available in USB mode is lost. In
addition, when apcupsd attempts to instruct the UPS to kill the power, it
begins cycling about 4 times a second between battery and line. The solution
to the problem (thanks to Tom Suzda) is to unplug the UPS and while it is
still chattering, press the power button (on the front of the unit) until the
unit beeps and the chattering stops. After that the UPS should behave normally
and power down 1-2 minutes after requested to do so.  

An amazing discovery by slither\_man allows one to build a CUSTOM-RJ45 cable
(documented above) and run the BackUPS CS (and probably also the ES) in Smart
mode. Running it this way provides all the same information that you would get
by running it in USB mode. As a consequence, we recommend that you either
purchase (where I don't know) or build your own CUSTOM-RJ45 cable rather than
use the 940-0128A cable.  

Thanks to all the people who have helped test this and have provided
information on the cable wiring, our best guess for the cable schematic is the
following: 

\footnotesize
\begin{verbatim}
     
     computer      --------- Inside the Connector---------  UPS
     DB9-F         |                                     |  RJ45
     pin - signal  |                                     |  Pin - Color
                   |                                     |
      4     DSR  ->|---+                                 |
                   |   |  diode   resistor               |
      6     DTR  ->|---+---->|----/\/\/\---o kill power  |  8  Orange
                   |                                     |
      1     DCD  <-|----+                                |
                   |    |                                |
      2     RxD  <-|----+----------------+--o low battery|  3  Brown
                   |                     |               |
      7     RTS  ->|----------+--/\/\/\--+               |
                   |          |                          |
                   |          +--/\/\/\--+               |
                   |                     |               |
      8     RI   <-|----+----------------+--o on battery |  2  Black
                   |    |                                |
      9     CTS  <-|----+                                |
                   |                         signal      |
      5     GND  --|-----------------------o ground      |  7  Red
                   |                                     |
      3     TxD    |                                     |
                   |                         chassis     |
      Chassis/GND  |-----------------------o ground      |  4  Black
                   |                                     |
                   |          Not connected              |  1, 5, 6, 9, 10
                   --------------------------------------
     
     The RJ45 pins are: looking at the end of the connector:
     
     10 9 8 7 6 5 4 3 2 1
     _______________________
     | . . . . . . . . . . |
     |                     |
     -----------------------
            |____|
     
\end{verbatim}
\normalsize

From Jan Babinski jbabinsk at pulsarbeacon dot com: (many thanks)  

The BackUPS XS1000(BX-1000) ships with the 940-0127B usb cable and the
optional 940-0128D serial cable. I'm trying to get it running under OpenBSD so
usb is out of the question. I've tried smart serial with the custom cable, but
that's not working, so I'm stuck with dumb serial over 940-0128D.  

I have noticed that this cable is not supported by apcupsd so I set out to
find out how much different this cable is compared with the supported ones.
After tearing the cable apart (see schematic below) I noticed that it is
functionally similar to the 940-0128A cable except for NC on (6) DTR and (2)
RD on the computer side. 

\footnotesize
\begin{verbatim}
     Cable 940-0128D
     
     DB9(Computer)               RJ45-10(UPS)
     
      (5)     (1)                 ____________
     ( o o o o o )               [ oooooooooo ]
      \ o o o o /                [____________]
       (9)   (6)                 (10)  [_]  (1)
     
     
      RI(9)<---+
               |
     CTS(8)<---+--- E   2N2222(NPN)
                     \|___
                ____ /| B |
               |    C     |
               |          |
               +---vvvv---+--[>|------<(2)OnBatt
     RTS(7)>---|    2k      1N5819
               +---vvvv---+--[>|------<(3)LowBatt
               |          |
               +--- C     |
                     \|___|
                     /| B
     DCD(1)<------- E    2N2222(NPN)
     
     DTR(4)>-------------------------->(8)KillPwr
     
     GND(5)----------------------------(7)Signal GND
     (Shield)--------------------------(4)Chassis GND
     
     
     Cable 940-0127B
     
     USB(Computer)      RJ45-10(UPS)
      _________          ____________
     | = = = = |        [ oooooooooo ]
     |_________|        [____________]
      (1)   (4)         (10)  [_]  (1)
     
       +5V(1)-----------(1)+5V
     DATA+(2)-----------(9)DATA+
     DATA-(3)-----------(10)DATA-
       GND(4)-----------(7)Signal GND
     (Shield)-----------(4)Chassis GRND
     
     
\end{verbatim}
\normalsize

\label{Win32-Implementation-Restrictions-for-Simple-UPSes}

\subsection*{Win32 Implementation Restrictions for Simple UPSes}
\index{UPSes!Win32 Implementation Restrictions for Simple }
\index{Win32 Implementation Restrictions for Simple UPSes }
\addcontentsline{toc}{subsection}{Win32 Implementation Restrictions for
Simple UPSes}

\label{index-Cables-213}
\label{index-Windows-214}
Due to inadequacies in the Win32 API, it is not possible to set/clear/get all
the serial port line signals. apcupsd can detect: CTS, DSR, RNG, and CD. It
can set and clear: RTS and DTR.  

This imposes a few minor restrictions on the functionality of some of the
cables. In particular, LineDown on the Custom Simple cable, and Low Battery on
the 0023A cable are not implemented. 

\label{Internal-Apcupsd-Actions-for-Simple-Cables}

\subsection*{Internal Apcupsd Actions for Simple Cables}
\index{Internal Apcupsd Actions for Simple Cables }
\index{Cables!Internal Apcupsd Actions for Simple }
\addcontentsline{toc}{subsection}{Internal Apcupsd Actions for Simple
Cables}

\label{index-Cables-215}

\footnotesize
\begin{verbatim}
     
     This section describes how apcupsd 3.8.5 (March 2002)
     treats the serial port line signals for simple cables.
     
     apcaction.c:
      condition = power failure detected
      cable = CUSTOM_SIMPLE
      action = ioctl(TIOCMBIS, DTR)      set DTR (enable power bit?)
     
     apcaction.c:
      condition = power back
      cable = CUSTOM_SIMPLE
      action = ioctl(TIOCMBIC, DTR)      clear DTR (clear power bit)
      action = ioctl(TIOCMBIC, ST)       clear ST (TxD)
     
     apcserial.c:
      condition = serial port initialization
      cable = 0095A, 0095B, 0095C
      action = ioctl(TIOMBIC, RTS)       clear RTS (set PnP mode)
     
      cable = 0119A, 0127A, 0128A
      action = ioctl(TIOMBIC, DTR)       clear DTR (killpower)
      action = ioctl(TIOMBIS, RTS)       set   RTS (ready to receive)
     
     apcserial.c:
      condition = save_dumb_status
      cable = CUSTOM_SIMPLE
      action = ioctl(TIOMBIC, DTR)       clear DTR (power bit?)
      action = ioctl(TIOMBIC, RTS)       clear RTS (killpower)
     
      cable = 0020B, 0020C, 0119A, 0127A, 0128A
      action = ioctl(TIOMBIC, DTR)       clear DTR (killpower)
     
      cable = 0095A, 0095B, 0095C
      action = ioctl(TIOMBIC, RTS)       clear RTS (killpower)
      action = ioctl(TIOMBIC, CD)        clear DCD (low batt)
      action = ioctl(TIOMBIC, RTS)       clear RTS (killpower) a second time!
     
     apcserial.c:
      condition = check_serial
     
      cable = CUSTOM_SIMPLE
      action = OnBatt = CD
      action = BattLow = CTS
      action = LineDown = SR
     
      cable = 0020B, 0020C, 0119A, 0127A, 0128A
      action = OnBatt = CTS
      action = BattLow = CD
      action = LineDown = 0
     
      cable = 0023A
      action = Onbatt = CD
      action = BattLow = SR
      action = LineDown = 0
     
      cable = 0095A, 0095B, 0095C
      action = OnBatt = RNG
      action = BattLow = CD
      action = LineDown = 0
     
     
     apcserial.c
      condition = killpower
     
      cable = CUSTOM_SIMPLE, 0095A, 0095B, 0095C
      action = ioctl(TIOMCBIS, RTS)      set RTS (kills power)
      action = ioctl(TIOMCBIS, ST)       set TxD
     
      cable = 0020B, 020C, 0119A, 0127A, 0128A
      action = ioctl(TIOMCBIS, DTR)      set DTR (kills power)
     
\end{verbatim}
\normalsize

\label{RS232-Wiring-and-Signal-Conventions}

\subsection*{RS232 Wiring and Signal Conventions}
\index{Conventions!RS232 Wiring and Signal }
\index{RS232 Wiring and Signal Conventions }
\addcontentsline{toc}{subsection}{RS232 Wiring and Signal Conventions}

\label{index-Cables-216}

\addcontentsline{lot}{table}{RS232 Wiring and Signal Conventions}
\begin{longtable}{llll}
{DB-25 Pin \#} & {DB-9 Pin \#} & {Name} & {DTE-DCE Description 
 } \\
{1} & {{---}} & {FG} & {{--} Frame Ground/Chassis GND 
 } \\
{2} & {3} & {TD} & {{--}\gt{} Transmitted Data, TxD 
 } \\
{3} & {2} & {RD} & {\lt{}{--} Received Data, RxD 
 } \\
{4} & {7} & {RTS} & {{--}\gt{} Request To Send 
 } \\
{5} & {8} & {CTS} & {\lt{}{--} Clear To Send 
 } \\
{6} & {6} & {DSR} & {\lt{}{--} Data Set Ready 
 } \\
{7} & {5} & {SG} & {{--}{---} Signal Ground, GND 
 } \\
{8} & {1} & {DCD} & {\lt{}{--} Data Carrier Detect 
 } \\
{9} & {{---}} & {{---}} & {{--} Positive DC test voltage 
 } \\
{10} & {{---}} & {{---}} & {{--} Negative DC test voltage 
 } \\
{11} & {{---}} & {QM} & {\lt{}{--} Equalizer mode 
 } \\
{12} & {{---}} & {SDCD} & {\lt{}{--} Secondary Data Carrier Detect 
 } \\
{13} & {{---}} & {SCTS} & {\lt{}{--} Secondary Clear To Send 
 } \\
{14} & {{---}} & {STD} & {{--}\gt{} Secondary Transmitted Data 
 } \\
{15} & {{---}} & {TC} & {\lt{}{--} Transmitter (signal) Clock 
 } \\
{16} & {{---}} & {SRD} & {\lt{}{--} Secondary Receiver Clock 
 } \\
{17} & {{---}} & {RC} & {{--}\gt{} Receiver (signal) Clock 
 } \\
{18} & {{---}} & {DCR} & {\lt{}{--} Divided Clock Receiver 
 } \\
{19} & {{---}} & {SRTS} & {{--}\gt{} Secondary Request To Send 
 } \\
{20} & {4} & {DTR} & {{--}\gt{} Data Terminal Ready 
 } \\
{21} & {{---}} & {SQ} & {\lt{}{--} Signal Quality Detect 
 } \\
{22} & {9} & {RI} & {\lt{}{--} Ring Indicator 
 } \\
{23} & {{---}} & {{---}} & {{--}\gt{} Data rate selector 
 } \\
{24} & {{---}} & {{---}} & {\lt{}{--} Data rate selector 
 } \\
{25} & {{---}} & {TC} & {\lt{}{--} Transmitted Clock  
}

\end{longtable}

\label{Pin-Assignment-for-the-Serial-Port-_005bRS_002d232C_005d_003b-25_002dpi%
n-and-9_002dpin_003b-Female-End}

\subsection*{Pin Assignment for the Serial Port (RS-232C), 25-pin and
9-pin, Female End}
\index{Pin Assignment for the Serial Port (RS-232C), 25-pin and 9-pin, Female
End }
\index{End!Pin Assignment for the Serial Port RS-232C 25-pin and 9-pin Female
}
\addcontentsline{toc}{subsection}{Pin Assignment for the Serial Port
(RS-232C), 25-pin and 9-pin, Female End}

\label{index-Cables-217}

\footnotesize
\begin{verbatim}
     
        13                         1         5         1
      _______________________________      _______________
      \  . . . . . . . . . . . . .  /      \  . . . . .  /    RS232-connectors
       \  . . . . . . . . . . . .  /        \  . . . .  /     looking into the
        ---------------------------          -----------      end of the cable.
        25                      14            9       6
     
     The diagram above represents the Female end of the cable. The
     male end is the same, but looking from inside the cable.
     
      DTE : Data Terminal Equipment (i.e. computer)
      DCE : Data Communications Equipment (i.e. UPS)
      RxD : Data received; 1 is transmitted "low", 0 as "high"
      TxD : Data sent; 1 is transmitted "low", 0 as "high"
      DTR : DTE announces that it is powered up and ready to communicate
      DSR : DCE announces that it is ready to communicate; low=modem hang-up
      RTS : DTE asks DCE for permission to send data
      CTS : DCE agrees on RTS
      RI  : DCE signals the DTE that an establishment of a connection is attempted
      DCD : DCE announces that a connection is established
\end{verbatim}
\normalsize

\label{Ioctl-to-RS232-Correspondence}

\subsection*{Ioctl to RS232 Correspondence}
\index{Ioctl to RS232 Correspondence }
\index{Correspondence!Ioctl to RS232 }
\addcontentsline{toc}{subsection}{Ioctl to RS232 Correspondence}

\label{index-Cables-218}

\footnotesize
\begin{verbatim}
     
     #define TIOCM_LE        0x001
     #define TIOCM_DTR       0x002
     #define TIOCM_RTS       0x004
     #define TIOCM_ST        0x008
     #define TIOCM_SR        0x010
     #define TIOCM_CTS       0x020
     #define TIOCM_CAR       0x040
     #define TIOCM_RNG       0x080
     #define TIOCM_DSR       0x100
     #define TIOCM_CD        TIOCM_CAR
     #define TIOCM_RI        TIOCM_RNG
     #define TIOCM_OUT1      0x2000
     #define TIOCM_OUT2      0x4000
\end{verbatim}
\normalsize
